Density Matrix Susceptibility

### 2018-05-25 papers/1_dmatpol/paper-r1.tex
Comparison with distributed polarizability model added

One very important feature of the susceptibility from Eq.~\eqref{e:final-model.HF} is that
the density matrix increment dependent {\emph only} by susceptibility element associated with the same
pair of AO's. Therefore, there is no mixing in the AO's, rather the increment is proportional to the 
vector susceptibility, different for each density matrix element.

@@

In order to validate all the models discussed in this Contribution,
the electric field\hyp{}induced interaction energy, 
the dipole moment and the quadrupole moment were analysed in detail.
The electric field\hyp{}induced interaction energy at the HF level, $\delta E$,
was computed from the following exact formula:
%
\begin{multline}\label{e:dE}
 \delta E[{\bf F}] = 
                     \Tr{\left\{ \delta {\bf D} \cdot
                                \left( {\bf H}^{(0)} + {\bf f}^{(0)} \right) \right\}}
                   + \Tr{\left\{ \left( {\bf D}^{(0)} + \delta {\bf D} \right) \cdot
                                \left( 2{\bf V} + 2\delta {\bf J} - \delta {\bf K}\right) \right\}} 
                   + E_{\rm nuc}({\bf F}) \;.
\end{multline}
%
In Eq.~\eqref{e:dE}, ${\bf f}^{(0)}$ and ${\bf H}^{(0)}$ are the Fock matrix and 
the one\hyp{}electron Hamiltonian matrix of an unperturbed HF ground state,
${\bf V}$ is the one\hyp{}electron external electrostatic potential matrix whereas $\delta {\bf J}$
and $\delta {\bf K}$ are electric field\hyp{}induced changes in the Coulomb and the exchange HF matrices due to 
the polarization of the electron density distribution, respectively. $E_{\rm nuc}({\bf F})$ denotes the interaction energy
due to the nuclei.
In order to compare the performance of the distributed polarizability approach,
the approximate interaction energy was also evaluated by using the expression:
%
\begin{equation}\label{e:dE-dpol}
 \delta E[{\bf F}] \approx 2 \Tr{\left\{ {\bf D}^{(0)} {\bf V}\right\} } 
                   + \frac{1}{2} \sum_i^{\rm occ} {\BM \alpha}_i : {\bf F}({\bf r}_i) \otimes {\bf F}({\bf r}_i)
                   + E_{\rm nuc}({\bf F}) \;,
\end{equation}
%
in which ${\BM \alpha}_i$ is the distributed anisotropic polarizability tensor associated with the
$i$th localized molecular orbital (LMO) centroid, ${\bf r}_i = \tbraket{\phi_i}{\bf r}{\phi_i}$.
The distributed polarizabilities were computed by utilizing the coupled\hyp{}perturbed Hartree\hyp{}Fock
method. Orbital localization was performed by using the Boys method. Note that in Eq.~\eqref{e:dE-dpol}, 
multipole approximation is introduced only for the part associated with the electric field\hyp{}induced
change in the one\hyp{}particle density matrix, i.e., the polarization of the electron density distribution.

>>>>

One very important feature of the susceptibility from Eq.~\eqref{e:final-model.HF} is that
the density matrix increment is dependent \emph{only} on the susceptibility element associated with the same
pair of AO's. Therefore, there is no mixing in the AO's, rather the increment is proportional to the 
vector susceptibility, different for each density matrix element.

@@

In order to validate all the models discussed in this Contribution,
the electric field\hyp{}induced interaction energy, 
the dipole moment and the quadrupole moment were analysed in detail.
The electric field\hyp{}induced interaction energy at the HF level, $\delta E$,
was computed from the following exact formula:
%
\begin{multline}\label{e:dE}
 \delta E[{\bf F}] = 
                     \Tr{\left\{ \delta {\bf D} \cdot
                                \left( {\bf H}^{(0)} + {\bf f}^{(0)} \right) \right\}}
                   + \Tr{\left\{ \left( {\bf D}^{(0)} + \delta {\bf D} \right) \cdot
                                \left( 2{\bf V} + 2\delta {\bf J} - \delta {\bf K}\right) \right\}} 
                   + E_{\rm nuc}({\bf F}) \;.
\end{multline}
%
In Eq.~\eqref{e:dE}, ${\bf f}^{(0)}$ and ${\bf H}^{(0)}$ are the Fock matrix and 
the one\hyp{}electron Hamiltonian matrix of an unperturbed HF ground state,
${\bf V}$ is the one\hyp{}electron external electrostatic potential matrix whereas $\delta {\bf J}$
and $\delta {\bf K}$ are electric field\hyp{}induced changes in the Coulomb and the exchange HF matrices due to 
the polarization of the electron density distribution, respectively. $E_{\rm nuc}({\bf F})$ denotes the interaction energy
due to the nuclei.
In order to compare the performance of the distributed polarizability approach,
the approximate interaction energy was also evaluated by using the expression:
%
\begin{equation}\label{e:dE-dpol}
 \delta E[{\bf F}] \approx 2 \Tr{\left\{ {\bf D}^{(0)} {\bf V}\right\} } 
                   - \frac{1}{2} \sum_i^{\rm occ} {\BM \alpha}_i : {\bf F}({\bf r}_i) \otimes {\bf F}({\bf r}_i)
                   + E_{\rm nuc}({\bf F}) \;,
\end{equation}
%
in which ${\BM \alpha}_i$ is the distributed anisotropic polarizability tensor associated with the
$i$th localized molecular orbital (LMO) centroid, ${\bf r}_i = \tbraket{\phi_i}{\bf r}{\phi_i}$.
The distributed polarizabilities were computed by utilizing the coupled\hyp{}perturbed Hartree\hyp{}Fock
method. Orbital localization was performed by using the Boys method. Note that in Eq.~\eqref{e:dE-dpol}, 
multipole approximation is introduced only for the part associated with the electric field\hyp{}induced
change in the one\hyp{}particle density matrix, i.e., the polarization of the electron density distribution.
